---
title: "Basic EDA"
author: "R.M. Palmer"
date: "11/15/2019"
output: html_document
---

# Basic EDA

```{r eda_basic, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gridExtra)
library(grid)
library(ggplot2)
library(lattice)

```

## Check for NA values

```{r sanity_check}
# check for NA values
train_raw %>% 
  select_if(function(x) any(is.na(x))) %>% 
  summarise_each(funs(sum(is.na(.)))) -> raw_NA
raw_NA
```

## Investigate Rate variables

```{r crossplot_wage_rates}
train_raw %>% 
  select(MonthlyIncome,DailyRate,HourlyRate,MonthlyRate) %>% pairs()
```

The rate variables do not seem correlated with Monthly Income.
We will discard them and stick with only Monthly Income.

## Overall summary
```{r}
str(train_raw)
cat ('distance from home')
min(train_raw$DistanceFromHome)
cat ('Education')
min(train_raw$Education)
cat ('EnvironmentSatisfaction')
min(train_raw$EnvironmentSatisfaction)
cat ('JobInvolvement')
min(train_raw$JobInvolvement)
cat ('JobLevel')
min(train_raw$JobLevel)
cat ('JobSatisfaction')
min(train_raw$JobSatisfaction)
cat ('NumCompaniesWorked')
min(train_raw$NumCompaniesWorked)
cat ('PercentSalaryHike')
min(train_raw$PercentSalaryHike)
cat ('PerformanceRating')
min(train_raw$PerformanceRating)
cat ('RelationshipSatisfaction')
min(train_raw$RelationshipSatisfaction)
cat ('StockOptionLevel')
min(train_raw$StockOptionLevel)
cat ('TotalWorkingYears')
min(train_raw$TotalWorkingYears)
cat ('TrainingTimesLastYear')
min(train_raw$TrainingTimesLastYear)
cat ('WorkLifeBalance')
min(train_raw$WorkLifeBalance)
cat ('YearsAtCompany')
min(train_raw$YearsAtCompany)
cat ('YearsInCurrentRole')
min(train_raw$YearsInCurrentRole)
cat ('YearsSinceLastPromotion')
min(train_raw$YearsSinceLastPromotion)
cat ('YearsWithCurrManager')
min(train_raw$YearsWithCurrManager)

```

No variables are negative.  When applying log transform, we will add 1 to NumCompaiesWorked, StockOptionLevel, TotalWorkingYears, TrainingTimesLastYear, YearsAtCompany, YearsInCurrentRole, YearsSinceLastPromotion, and YearsWithCurrManager.